Laser tracker placement optimization for highly flexible manufacturing systems

### 2024-08-05 paper/paper.tex
added intro, some templating.

%\documentclass[a4paper, 10pt, conference]{ieeeconf}      % Use this line for a4 paper

\IEEEoverridecommandlockouts                              % This command is only needed if
                                                          % you want to use the \thanks command

\overrideIEEEmargins                                      % Needed to meet printer requirements.

%In case you encounter the following error:
%Error 1010 The PDF file may be corrupt (unable to open PDF file) OR
%Error 1000 An error occurred while parsing a contents stream. Unable to analyze the PDF file.
%This is a known problem with pdfLaTeX conversion filter. The file cannot be opened with acrobat reader
%Please use one of the alternatives below to circumvent this error by uncommenting one or the other
%\pdfobjcompresslevel=0
%\pdfminorversion=4

% See the \addtolength command later in the file to balance the column lengths
% on the last page of the document

@@

\title{\LARGE \bf
Laser Tracker Placement Optimization for highly flexible manufacturing systems
}

\author{Jan Baumgärtner$^{1*}$, Max Goebels$^{1*}$,   Alexander Puchta$^{1}$ and Jürgen Fleischer$^{1}$% <-this % stops a space
\thanks{*These authors contributed equally to this work.}% <-this % stops a space
\thanks{$^{1}$Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer are with the wbk Institute of Production Science,
        Karlsruhe Institute of Technology, 76131 Karlsruhe, Germany
        {\tt\small max.goebels@kit.edu}}%
}

\begin{document}

\maketitle
\thispagestyle{empty}
\pagestyle{empty}

\section*{Abstract}
In recent years, modular and reconfigurable manufacturing systems have gained attention for their adaptability. 
While system design is now highly automated, designing the verification setup remains a manual design problem involving tasks like placing external sensors to measure performance. 
This paper argues that design optimization approaches for manufacturing systems can also automate the design of the verification setup.
This is demonstrated using the placement of a laser tracker system with markers that due to their line-of-sight complexity and simultaneous measurement of multiple systems can be challenging to design.
We present a straightforward optimization approach to demonstrate potential automation in this area.
\section{Introduction}

\section{State of the art}
The work presented in this paper intersects two research fields: manufacturing systems optimization and sensor placement optimization. 
Our primary question is what aspects of a manufacturing system we need to verify. 
We focus on system-based verification rather than product-based, aiming to measure the performance of the manufacturing system, not the resulting product quality.
In particular we focuss on approaches that solves what~\cite{reconfigurable_production} calls reconfiguration design problem.
This involves either combining fixed modules to form a system or directly optimizing the sytems physical design.
 Examples include robot cell optimization problems~\cite{previous_work, stiffness_placement} and robot design optimization problems~\cite{task_synthesis, ad-hoc_manipulator, multi_objective}. 
 For a manufacturing perspective, works such as~\cite{johannes_1, johannes_2} are relevant.
 All of these approaches focus on phyiscal capabilities such as payload, accuracy, stiffness, force, weight, and motion.

@@

\section{Formulation of the optimization problem}
To solve the joint optimization problem of marker and laser tracker pose we first need to define the objective function.
Our goal is ofcourse to maximize the visibility of the marker to the laser tracker.
Similar to approaches such as~\cite{ieee_sensors} we divide the visibility into two components.
Line of sight visibility and field of view visibility.
Both  are depicted in figure~\ref{fig:visibility}.
\begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/visibility.png}
        \caption{Field of view (VOW) angle range and line of sight (LOS) components of marker visibility}
        \label{fig:visibility}
\end{figure}
Mathematically we can define line of sight visibility (los visibility) as the following binary function:
\begin{equation}
    f(p_m,p_l) = \begin{cases}
    -1 & \text{if } \text{marker is in line of sight of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
where $p_m$ is the position of the marker, $p_l$ is the position of the laser tracker.
Note that $p_m$ is typically dependent on time as it is attached to a moving manufacuting system such as a robot.
The field of view objective can similarly be defined as a binary function:
\begin{equation}
    g(p_m,p_l) = \begin{cases}
    -1 & \text{if } \text{marker is in the field of view of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
However in practice it might make sense to not be close to the boundary of the field of view.
Instead one can define a continuous function that is 1 in the center of the field of view and 0 at the boundary.
The simplest way to encode this using the angle $\alpha$ between the normal of the marker and the line of sight of the laser tracker.
\begin{equation}
    g(p_m,p_l) = -\cos(\alpha)
\end{equation}
Apart from line of sight we also want to optimize the measurment performance of the laser tracker.
The accuracy of the lasertracker is mainly dependent on the distance between the marker and the laser tracker.
The further away the marker is the less accurate the measurment will be. However there is also a minimum measurment distance that the laser tracker can handle.
To encode this we can define a penalty function that penalizes markers that are too far away from the laser tracker.
\begin{equation}
    p(p_m,p_l) =  \begin{cases}
        d & \text{if marker is in measurement range of the laser tracker} \\
        \rho & \text{otherwise}
    \end{cases}
\end{equation}
where $d$ is the distance between the marker and the laser tracker, and $\rho$ is a large penalty factor.
The final objective function is then a weighted sum of the two components:
\begin{equation}
    h(p_m,p_l) = \int f(p_m,p_l) +  g(p_m,p_l) + p(p_m,p_l) \, dt
    \label{eq:objective}
\end{equation}

@@

While $T_{m}$ is a 4x4 matrix, it can be parameterized using the marker's position ($p_{p_m}$) and orientation ($p_{o_m}$), resulting in a 6-dimensional design space for the marker pose.
To avoid unrealistic positions, we constrain this space to keep the marker close to the manufacturing system's surface.
This constraint can be represented cylindrically:
\begin{equation}
    \begin{split}
        \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} &\leq r \\
        0 \leq p_{p_m,z} &\leq h
    \end{split}
\end{equation}I

@@

\section{Optimization algorithm}
There are a large number of possible optimization algorithms that can be used to solve this problem.
Following the no free lunch theorem \cite{no_free_lunch_theorem} we want to pick the algorithm that makes the most use of the underlying structure of the problem.
In this case the problem is a continuous optimization problem with a relatively low number of design variables.
On the other hand it is a gradient free optimization problem due to the line of sight visibility function.
These two properties make the problem well suited for the particle swarm optimization (PSO) algorithm \cite{pso}.
Particle swarm optimization is a computational method for finding the optimal solution to a problem.
It is inspired by the behaviour of swarms in nature, such as flocks of birds or schools of fish, which exhibit emergent behaviour that is intelligent and efficient.
 In PSO, a group of particles (also called agents or individu-als) move through a search space, according to the fol-lowing update rules:
\begin{equation}
        \begin{split}
        x_{i,j} &= x_{i,j} + v_{i,j} \\
    v_{i,j} &= wv_{i,j} + c_1r_1(p_{i,j}-x_{i,j}) + c_2r_2(p_{g,j}-x_{i,j})
        \end{split}
\end{equation}
where $x_{i,j}$ is the position of particle $i$ in dimension $j$, $v_{i,j}$ is the velocity of particle $i$ in dimension $j$, $w$ is the inertia weight,
 $c_1$ and $c_2$ are the cognitive and social coefficients, $r_1$ and $r_2$ are random numbers between 0 and 1, $p_{i,j}$ is the best position of particle $i$ in dimension $j$ and $p_{g,j}$
is the best position of the swarm in dimension $j$.
In our case the particles are a 9 dimensional vector based on the position of the laser tracker and the pose of the marker.
The best particle is the one that maximizes the objective function defined in equation~\ref{eq:objective}.
The hyperparameters we need to tune are collected in table~\ref{tab:hyperparameters}.
Note that the pso algorithm has no direct way to encode the constraints we defined earlier.
Instead we can use a penalty function that penalizes any particles that violate the constraints.
This is a common approach in optimization and is also used in the implementation of the pso algorithm in the scipy library. %TODO citation needed
The penalty function can be defined as:
\begin{equation}
    p(x) =  \begin{cases}
        0 & \text{if } x \text{ satisfies all constraints} \\
        \rho & \text{otherwise}
    \end{cases}
\end{equation}
where $\rho$ is a large penalty factor.

@@

\section{Simulation Environment}
Evaluating the objective function requires a simulation of the manufacturing system.
The tricky part here is that since the manufacturing is actively manufacturing parts we not only need to simulate the behavior of the system itself but the changing shape of any parts that are being manufactured.
For this purpose we are using the pybullet-industrial package \cite{pybullet_industrial}.
This extension of the pybullet physics engine allows us to simulate not only the robot but also the manufacturing process.
For the purposes of this paper the markers are implemented as a pybullet industrial endeffector obect that is attached to the robot.
A image of the simulation environment is shown in figure \ref{fig:simulation}.
\begin{figure}
        \centering
        \includegraphics[width=\textwidth]{figures/simulation.png}
        \caption{Simulation environment used for the optimization}
        \label{fig:simulation}
\end{figure}

@@

\section{Acknowledgements}

%Bibliography
%TODO checken ob das stimmt in template
\printbibliography

>>>>

% to typeset URLs, URIs, and DOIs
\usepackage{url}
\def\UrlFont{\rmfamily}

@@

\begin{document}
\mainmatter              % start of a contribution
\title{
Laser tracker placement optimization for highly flexible manufacturing systems
}
\titlerunning{Laser tracker placement}

% latex magic to format multiple authors affiliations
\newcommand*\samethanks[1][\value{footnote}]{\footnotemark[#1]}
\makeatletter
% *, 2, 3, ...
\renewcommand*{\@fnsymbol}[1]{\ifcase#1\or\@arabic{#1}\else*\fi}
\makeatother

\author{Jan Baumgärtner \inst{1,2}, 
Max Goebels \inst{1,2},   Alexander Puchta \inst{1,2} and Jürgen Fleischer\inst{1}% <-this % stops a space
}
\institute{wbk Institute of Production Science, Karlsruhe Institute of Technology, 
76131 Karlsruhe, Germany \email{max.goebels@kit.edu}
\and
These authors contributed equally to this work.
}

\authorrunning{Baumgärtner et al.}
\tocauthor{Jan Baumgärtner, Max Goebels, Alexander Puchta, and Jürgen Fleischer}

\maketitle
%\thispagestyle{empty}
%\pagestyle{empty}

\begin{abstract}
In recent years, modular and reconfigurable manufacturing systems have gained attention for their adaptability. 
While system design is now highly automated, designing the verification setup remains a manual design problem involving tasks like placing external sensors to measure performance. 
This paper argues that design optimization approaches for manufacturing systems can also automate the design of the verification setup.
This is demonstrated using the placement of a laser tracker system with markers that due to their line-of-sight complexity and simultaneous measurement of multiple systems can be challenging to design.
We present a straightforward optimization approach to demonstrate potential automation in this area.
\keywords{flexible manufacturing systems, sensor placement optimization, laser tracker, particle swarm optimization}
\end{abstract}
\section{Introduction}

The global market for flexible manufacturing systems (FMS) is experiencing significant growth, driven by the increasing demand for customized products and shorter product life cycles~\cite{westkaemperEinfuehrungOrganisationProduktion2006}. FMS are characterized by their ability to adapt to changing requirements, such as variations in product design, production volume, and production sequence. Modular reconfigurable machine tools (MRMT) are a key component of FMS, as they can modify their physical structure to meet evolving product demands~\cite{Padayachee2012}.

The majority of research in the field of MRMT has concentrated on the development of architectures that adhere to open or semi-open standards, the modular components, and methods for optimizing configurations based on performance metrics~\cite{gadallaRecentAdvancesResearch2017, xuMethodDesignModular2017}.

Despite the aforementioned advancements, a significant gap in the research field remains: the verification of the performance of the aforementioned proposed configurations is often limited to simulations, which are susceptible to a reality gap. Alternatively, manual experimental setups using external sensors are used. However, these manual setups are not only time-consuming but also require expert knowledge and are susceptible to errors, rendering them inadequate for rigorous verification.

To address this issue, our paper introduces an automated approach to task-based sensor placement optimization for the verification of MRMT. For the purposes of this study, we consider a robot cell to be an ideal test case for the proposed approach, providing an upper bound for the complexity of task trajectories. This is based on the recognition that although the reconfiguration of robot cells is less pronounced, the task trajectories within them are typically more complex than those typically found in MRMT applications. 
\section{State of the art}
The work presented in this paper intersects two research fields: manufacturing systems optimization and sensor placement optimization. 
Our primary question is what aspects of a manufacturing system we need to verify. 
We focus on system-based verification rather than product-based, aiming to measure the performance of the manufacturing system, not the resulting product quality.
In particular we focuss on approaches that solves what~\cite{reconfigurable_production} calls reconfiguration design problem.
This involves either combining fixed modules to form a system or directly optimizing the systems physical design. 
Examples include robot cell optimization problems~\cite{previous_work, stiffness_placement} and robot design optimization problems~\cite{task_synthesis, ad-hoc_manipulator, multi_objective}.  
For a manufacturing perspective, works such as~\cite{johannes_1, johannes_2} are relevant. 
All of these approaches focus on phyiscal capabilities such as payload, accuracy, stiffness, force, weight, and motion.

@@

\section{Formulation of the optimization problem}
To solve the joint optimization problem of marker and laser tracker pose we first need to define the objective function.
Our goal is ofcourse to maximize the visibility of the marker to the laser tracker.
Similar to approaches such as~\cite{ieee_sensors} we divide the visibility into two components.
Line of sight visibility and field of view visibility.
Both  are depicted in figure~\ref{fig:visibility}.
\begin{figure}
        \centering
        \includegraphics[width=0.8\textwidth]{figures/visibility.png}
        \caption{Field of view (FOV) angle range and line of sight (LOS) components of marker visibility}
        \label{fig:visibility}
\end{figure}
Mathematically we can define line of sight visibility (los visibility) as the following binary function:
\begin{equation}
    f(p_m,p_l) = \begin{cases}
    -1 & \text{if } \text{marker is in line of sight of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
where $p_m$ is the position of the marker, $p_l$ is the position of the laser tracker.
Note that $p_m$ is typically dependent on time as it is attached to a moving manufacturing system such as a robot.
The field of view objective can similarly be defined as a binary function:
\begin{equation}
    g(p_m,p_l) = \begin{cases}
    -1 & \text{if } \text{marker is in the field of view of the laser tracker} \\
    0 & \text{otherwise}
    \end{cases}
\end{equation}
However in practice it might make sense to not be close to the boundary of the field of view.
Instead one can define a continuous function that is 1 in the center of the field of view and 0 at the boundary.
The simplest way to encode this using the angle $\alpha$ between the normal of the marker and the line of sight of the laser tracker.
\begin{equation}
    g(p_m,p_l) = -\cos(\alpha)
\end{equation}
Apart from line of sight we also want to optimize the measurment performance of the laser tracker.
The accuracy of the lasertracker is mainly dependent on the distance between the marker and the laser tracker.
The further away the marker is the less accurate the measurment will be. However there is also a minimum measurment distance that the laser tracker requires to find a target.
To encode this we can define a penalty function that penalizes markers that are too far away from the laser tracker.
\begin{equation}
    p(p_m,p_l) =  \begin{cases}
        d & \text{if marker is in measurement range of the laser tracker} \\
        \rho & \text{otherwise}
    \end{cases}
\end{equation}
where $d$ is the distance between the marker and the laser tracker, and $\rho$ is a large penalty factor.
The final objective function is then a weighted sum of the two components:
\begin{equation}
    h(p_m,p_l) = \int f(p_m,p_l) +  g(p_m,p_l) + p(p_m,p_l) \, dt
    \label{eq:objective}
\end{equation}

@@

While $T_{m}$ is a 4x4 matrix, it can be parameterized using the marker's position ($p_{p_m}$) and orientation ($p_{o_m}$), resulting in a 6-dimensional design space for the marker pose.
To avoid unrealistic positions, we constrain this space to keep the marker close to the manufacturing system's surface.
This constraint can be represented cylindrically:
\begin{equation}
    \begin{split}
        \sqrt{p_{p_m,x}^2 + p_{p_m,y}^2} &\leq r \\
        0 \leq p_{p_m,z} &\leq h
    \end{split}
\end{equation}

@@

\section{Optimization algorithm}
There are a large number of possible optimization algorithms that can be used to solve this problem.
Following the no free lunch theorem \cite{no_free_lunch_theorem} we want to pick the algorithm that makes the most use of the underlying structure of the problem.
In this case the problem is a continuous optimization problem with a relatively low number of design variables.
On the other hand it is a gradient free optimization problem due to the line of sight visibility function.
These two properties make the problem well suited for the particle swarm optimization (PSO) algorithm \cite{pso}.
Particle swarm optimization is a computational method for finding the optimal solution to a problem.
It is inspired by the behaviour of swarms in nature, such as flocks of birds or schools of fish, which exhibit emergent behaviour that is intelligent and efficient.
 In PSO, a group of particles (also called agents or individu-als) move through a search space, according to the fol-lowing update rules:
\begin{equation}
        \begin{split}
        x_{i,j} &= x_{i,j} + v_{i,j} \\
    v_{i,j} &= wv_{i,j} + c_1r_1(p_{i,j}-x_{i,j}) + c_2r_2(p_{g,j}-x_{i,j})
        \end{split}
\end{equation}
where $x_{i,j}$ is the position of particle $i$ in dimension $j$, $v_{i,j}$ is the velocity of particle $i$ in dimension $j$, $w$ is the inertia weight,
 $c_1$ and $c_2$ are the cognitive and social coefficients, $r_1$ and $r_2$ are random numbers between 0 and 1, $p_{i,j}$ is the best position of particle $i$ in dimension $j$ and $p_{g,j}$
is the best position of the swarm in dimension $j$.
In our case the particles are a 9 dimensional vector based on the position of the laser tracker and the pose of the marker.
The best particle is the one that maximizes the objective function defined in equation~\ref{eq:objective}.
The hyperparameters we need to tune are collected in table~\ref{tab:hyperparameters}.
Note that the pso algorithm has no direct way to encode the constraints we defined earlier.
Instead we can use a penalty function that penalizes any particles that violate the constraints.
This is a common approach in optimization and is also used in the implementation of the pso algorithm in the scipy library~\cite{2020SciPy-NMeth}.
The penalty function can be defined as:
\begin{equation}
    p(x) =  \begin{cases}
        0 & \text{if } x \text{ satisfies all constraints} \\
        \rho & \text{otherwise}
    \end{cases}
\end{equation}
where $\rho$ is a large penalty factor.

@@

\section{Simulation Environment}
Evaluating the objective function requires a simulation of the manufacturing system.
The tricky part here is that since the manufacturing is actively manufacturing parts we not only need to simulate the behavior of the system itself but the changing shape of any parts that are being manufactured.
For this purpose we are using the pybullet-industrial package \cite{pybullet_industrial}.
This extension of the pybullet physics engine allows us to simulate not only the robot but also the manufacturing process.
For the purposes of this paper the markers are implemented as a pybullet industrial endeffector object that is attached to the robot.
A image of the simulation environment is shown in figure \ref{fig:simulation}.
\begin{figure}
        \centering
        \includegraphics[width=\textwidth]{figures/simulation.png}
        \caption{Simulation environment used for the optimization}
        \label{fig:simulation}
\end{figure}

@@

\section{Acknowledgements}
The authors declare that they have no known competing financial interests or personal relationships that could have appeared to influence the work reported in this paper.
This work was supported by the Ministry of Science, Research and Arts of the Federal State of Baden-Württemberg within the InnovationsCampus Future Mobility. The APC was funded by Innovationscampus Mobilität der Zukunft (ICM), grant number SdManu7.
The robot cell in the sample test case are part of the project funded by the Deutsche Forschungsgemeinschaft (DFG, German Research Foundation) - SFB-1574 - 471687386.
%Bibliography
%TODO checken ob das stimmt in template
\printbibliography
